---
title: "PPOL564"
output: 
  html_document:
    css: style.css
    theme: cerulean
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r,include=F}
# Packages for reading in the tables
require(tidyverse)
require(kableExtra)
```



<center><h1>Data Science I: Foundations</h1></center>
<center>**Fall 2019**</center>
<center>**Georgetown University**</center>

<br>



# Course Outline {.tabset .tabset-fade .tabset-pills}

## Course Schedule

<br>
<br>


```{r,echo=F,message=F,results='asis'}
read_csv('syllabus/fall-2019-timeline.csv') %>%
  select(-`Class No.`,-CD,-Assignment) %>% 
  filter(Topic != "No Class") %>% 
  {.[is.na(.)]="";.} %>% 
  kable(., "html", booktabs = T,align = c('c','l','l')) %>%
  # kable(., "latex", booktabs = T,align = c('c','l','l')) %>%
  kable_styling(latex_options = c("striped", 
                                  "scale_down","repeat_header"),
                full_width = T,font_size = 16) %>% 
  column_spec(1,width = '.5in') %>% 
  column_spec(2,width = '1in') %>% 
  column_spec(3,width = '4in') 
```

<br>
<br>
<br>
<br>


## Syllabus

<br>
<br>

- Access an [**Online Version**](syllabus/PPOL564-Syllabus.html)
- [**Download**](syllabus/PPOL564-Syllabus.pdf) a `.pdf` version locally. 

<br>
<br>
<br>
<br>

## Contact

<br>

#### [Eric Dunford](www.ericdunford.com) (Professor)

- **Office**: 404 Old North
- **Office Hours**: Tuesdays 3pm to 5pm
- **Email**: [eric.dunford@georgetown.edu](mailto:eric.dunford@georgetown.edu)

<br>

#### Andrea Quevedo (Teaching Assistant)

- **Office Hours**: by appointment
- **Email**: [aq38@georgetown.edu](mailto:aq38@georgetown.edu)

<br>
<br>
<br>
<br>

## Lab Schedule

<br>
<br>

To be announced.

<br>
<br>
<br>
<br>


# Lecture Materials

<br>
<br>

### Reproducibility

- [Lecture Slides](lectures/lecture_01/lecture-01.html)

### Version Control
### Python Notebooks
### Data Types in Python
### Control Sequences, Iteration, and Functions
### Comprehensions and Generators
### Numpy
### Data Wrangling with Pandas (part 1)
### Data Wrangling with Pandas (part 2)
### Exploratory Data Analysis
### Vectors 
### Trigonometry of Vectors 
### Matrix Transformations
### Matrix Operations and Inversions 
### Linear Regression 
### Eigen Decompositions
### Decompositions in Practice 
### Differentiation
### Optimizing Univariate Functions
### Optimizing Multivariate Functions
### Gradient Decent
### Constrained Optimization and Regularization
### Probability 
### Bayes Rule & Naive Bayes Algorithm
### Simulation and MCMC Sampling

<br>
---




