---
title: "PPOL564"
output: 
  html_document:
    css: style.css
    theme: cerulean
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r,include=F}
# Packages for reading in the tables
require(tidyverse)
require(kableExtra)
```



<center><h1>Data Science I: Foundations</h1></center>
<center>**Fall 2019**</center>
<center>**Georgetown University**</center>

<br>



# Course Outline {.tabset .tabset-fade .tabset-pills}

## Course Schedule

<br>
<br>


```{r,echo=F,message=F,results='asis'}
read_csv('syllabus/fall-2019-timeline.csv') %>%
  select(-`Class No.`,-CD,-Assignment) %>% 
  filter(Topic != "No Class") %>% 
  {.[is.na(.)]="";.} %>% 
  kable(., "html", booktabs = T,align = c('c','l','l')) %>%
  # kable(., "latex", booktabs = T,align = c('c','l','l')) %>%
  kable_styling(latex_options = c("striped", 
                                  "scale_down","repeat_header"),
                full_width = T,font_size = 16) %>% 
  column_spec(1,width = '.5in') %>% 
  column_spec(2,width = '1in') %>% 
  column_spec(3,width = '4in') 
```

<br>
<br>
<br>
<br>


## Syllabus

<br>
<br>

- Access an [**Online Version**](syllabus/PPOL564-Syllabus.html)
- [**Download**](syllabus/PPOL564-Syllabus.pdf) a `.pdf` version locally. 

<br>
<br>
<br>
<br>

## Readings

<br>
<br>

List of all [required reading](readings/reading-list.html).

<br>
<br>
<br>
<br>

## Recitation

<br>
<br>

Recitation meets most weeks on Mondays from 5:00 to 5:50 in **Car Bar 203**. Below is the schedule for when there is a lab. 

```{r,echo=F}
tribble(
  ~'Lab No.', ~Date, ~` `,
    1, "09-09", "Lab",
    2, "09-16", "Lab",
    3, "09-23", "Makeup Class (class does not meet earlier in the day)",
    "","09-30", "Attend the Data Science in Action Seminar",
    4, "10-07", "Lab",
    "", "10-14", "No Lab - Mid-Semester Holiday",
    "", "10-21", "Attend the Data Science in Action Seminar",
    5, "10-28", "Lab",
    6, "11-04", "Lab",
    "","11-11", "Attend the Data Science in Action Seminar",
    7, "11-18", "Lab",
    8, "11-25", "Lab",
    9, "12-02", "Lab",
    10, "12-09", "Lab"
) %>% 
  kable(., "html", booktabs = T,align = c('c','l','l')) %>%
  # kable(., "latex", booktabs = T,align = c('c','l','l')) %>%
  kable_styling(latex_options = c("striped", 
                                  "scale_down","repeat_header"),
                full_width = T,font_size = 16) %>% 
  column_spec(1,width = '1in') %>% 
  column_spec(2,width = '1in') %>% 
  column_spec(3,width = '4in') 
```


<br>
<br>
<br>
<br>

## Contact

<br>

#### [Eric Dunford](www.ericdunford.com) (Professor)

- **Office**: 404 Old North
- **Office Hours**: Tuesdays 3pm to 5pm
- **Email**: [eric.dunford@georgetown.edu](mailto:eric.dunford@georgetown.edu)

<br>

#### Andrea Quevedo (Teaching Assistant)

- **Office Hours**: by appointment
- **Email**: [aq38@georgetown.edu](mailto:aq38@georgetown.edu)

<br>
<br>
<br>
<br>

# Lecture Materials

<br>
<br>

### Reproducibility

- [**Lecture Slides**](lectures/lecture_01/lecture-01.html)

<!-- ### Version Control -->
<!-- ### Python Notebooks -->
<!-- ### Data Types in Python -->
<!-- ### Control Sequences, Iteration, and Functions -->
<!-- ### Comprehensions and Generators -->
<!-- ### Numpy -->
<!-- ### Data Wrangling with Pandas (part 1) -->
<!-- ### Data Wrangling with Pandas (part 2) -->
<!-- ### Exploratory Data Analysis -->
<!-- ### Vectors  -->
<!-- ### Trigonometry of Vectors  -->
<!-- ### Matrix Transformations -->
<!-- ### Matrix Operations and Inversions  -->
<!-- ### Linear Regression  -->
<!-- ### Eigen Decompositions -->
<!-- ### Decompositions in Practice  -->
<!-- ### Differentiation -->
<!-- ### Optimizing Univariate Functions -->
<!-- ### Optimizing Multivariate Functions -->
<!-- ### Gradient Decent -->
<!-- ### Constrained Optimization and Regularization -->
<!-- ### Probability  -->
<!-- ### Bayes Rule & Naive Bayes Algorithm -->
<!-- ### Simulation and MCMC Sampling -->

<br>
---




